---
title: "Ascidians"
author: "Shannon Eckhardt"
date: "2023-06-04"
output: html_document
---

```{r}
rm(list=ls())
library(tidyverse)
library(dplyr)
library(readxl)
```

Import data
```{r}
benthic.data <- read_excel("Benthic_cover.xlsx")
```

Extract ascidian data
```{r}
ascidians.full <- benthic.data[, c("Image ID", "Image name", "TP", "Site", "Transect_ID", "Transect_no", "ASCD")]
```

Convert TP to date format
```{r}
month_year <- strsplit(ascidians.full$TP, "_")
month <- sapply(month_year, function(x) x[1])
year <- sapply(month_year, function(x) x[2])
months <- c("January", "February", "March", "April", "May", "June", "July",
            "August", "Sept", "October", "November", "Dec") 
            # Sept and Dec because that's how they're saved in ascidians.full
month_numeric <- match(month, months)
year_numeric <- as.numeric(year)
date <- as.Date(paste(year_numeric, month_numeric, "01", sep = "-"))

# Save the date in a new column, other than TP
ascidians.full$Date <- date
```


Group by time point and site
```{r}
# Summed up cover = how many is there
ascidians.sum <- ascidians.full %>%
  group_by(TP, Site, Date) %>%
  summarize(sum_cover = sum(ASCD))

# remove NAs first
ascidians.sum <- na.omit(ascidians.sum)

# order the TPs
ascidians.sum <- ascidians.sum %>%
  mutate(TP = factor(TP, levels = c("Dec_2020", "April_2021", "April_2022", "Sept_2022", "Dec_2022",  "April_2023"))) %>%
  arrange(TP)

# simple plot
ggplot(ascidians.sum, aes(x = TP, y = sum_cover, fill = Site)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(x = "Time point", y = "Count", fill = "Site") +
  ggtitle("Count of ascidians over time and per site") +
  theme_classic()

```
***PERCENT COVER***
Percent cover per transect and per site
Boxplots with SE
Dotplots with SE
```{r}
# pivot table
t.cover <- benthic.data %>% pivot_longer(c(Acanthastrea:Valspp), names_to = "Genera", values_to = "Cover")

# calculate mean cover & SD and SE by transect
mean.cover <- t.cover %>%
  group_by(TP, Site, Transect_ID, Genera) %>%
  summarize(mean_cover = mean(Cover)) %>%
  mutate(SE = sd(mean_cover) / sqrt(n()),
         SD = sd(mean_cover))

# calculate sum of cover values for each transect - rowSums
# make new column called percent_cover - multiply by 100 to get percent
transect.percent.cover <- mean.cover %>%
  group_by(TP, Site, Transect_ID) %>%
  summarize(percent_cover = sum(mean_cover[Genera == "ASCD"]) / sum(mean_cover) * 100, .groups = "drop") %>%
  mutate(SE = sd(percent_cover)/sqrt(n()))

# percent cover per site
site.percent.cover <- mean.cover %>%
  group_by(TP, Site) %>%
  summarize(percent_cover = sum(mean_cover[Genera == "ASCD"]) / sum(mean_cover) * 100, .groups = "drop") %>%
  mutate(SE = sd(percent_cover)/sqrt(n()))

# percent cover per time point
tp.percent.cover <- mean.cover %>%
  group_by(TP) %>%
  summarize(percent_cover = sum(mean_cover[Genera == "ASCD"]) / sum(mean_cover) * 100, .groups = "drop") %>%
  mutate(SE = sd(percent_cover)/sqrt(n()))

# order the TPs
transect.percent.cover <- transect.percent.cover %>%
  mutate(TP = factor(TP, levels = c("Dec_2020", "April_2021", "April_2022", "Sept_2022", "Dec_2022",  "April_2023"))) %>%
  arrange(TP)

site.percent.cover <- site.percent.cover %>%
  mutate(TP = factor(TP, levels = c("Dec_2020", "April_2021", "April_2022", "Sept_2022", "Dec_2022",  "April_2023"))) %>%
  arrange(TP)

```

***Boxplot***
no errorbars on boxplot
```{r}
str(transect.percent.cover)
ggplot(transect.percent.cover , aes(x = TP, y = percent_cover, fill = Site)) +
  geom_boxplot() +
  labs(x = "Time Points", y = "Percent Cover (%)") +
  scale_fill_discrete(name = "Site") +
  theme_classic()
```

***Dotplot***
```{r}
str(site.percent.cover)
# plot per time point -> not the correct plot -> need to integrate TP and possibly site?
ggplot(site.percent.cover, aes(x = TP, y = percent_cover, colour = Site)) +
  geom_point(position=position_dodge(.9)) +
  geom_errorbar(aes(ymin = percent_cover - SE, ymax = percent_cover + SE),
                width = 0.4, size = 0.7,
                 position=position_dodge(.9)) +
  labs(x = "Time Points", y = "Cover (%)") +
  scale_fill_discrete(name = "Site") +
  scale_colour_discrete(name = "Site") +
  ylim(0,2.5) +
  theme_classic()
```

***FREQUENCY OF OCCURRENCE***
Calculate frequency of occurrence (presented as table with TP, Site and FOC) = proportion of transects that have them present (presence/absence) per TP = how widespread it could be in the bay
  For each site EB,SB, CB (then maybe group SB and EB as lagoon)
  
% frequency of ascidians = (# of quadrats where ascidians occur / total # of quadrats examined)* 100
```{r}
# first average benthic data at transect level (average over 10 quadrats)
mean.cover.transects <- t.cover %>%
  group_by(TP, Site, Transect_ID, Genera) %>%
  summarize(mean_cover = mean(Cover))

# pivot long table back to wide
mean.cover.transects <- mean.cover.transects %>%
  pivot_wider(names_from = Genera, values_from = mean_cover)

ascd.occurr <- mean.cover.transects %>%
  mutate(ascd_occurrence = ifelse(ASCD != 0, "present", "absent"))

# pick out TP, Sites and ascd_occurrence


sum(ascd.occurr$ascd_occurrence == "present") # only 8 transects where ascidians occur

# calculate FOC for each TRANSECT_ID
transect.foc <- ascd.occurr %>%
  group_by(TP, Site, Transect_ID) %>%
  summarize(foc = (sum(ascd_occurrence == "present") / length(ascd.occurr$Transect_ID))*100)
# not sure if this length(benthic.data$Transect_no) is correct

# calculate FOC for each Site
site.foc <- ascd.occurr %>%
  group_by(TP, Site) %>%
  summarize(foc = (sum(ascd_occurrence == "present") / length(ascd.occurr$Transect_ID))*100)

# calculate FOC per TP
TP.foc <- ascd.occurr %>%
  group_by(TP) %>%
  summarize(foc = (sum(ascd_occurrence == "present") / length(ascd.occurr$Transect_ID))*100)
```

***Relative cover***
- take out all hard coral substrate and not sand
- for anything that isn't coral or sand, what proportion is made up of ascidians
- how much of 100% is ascidian cover
- sum of ascidians divided by sum all other cover (including ascidians) minus hard coral and sand
- are they making up more cover than algae in some places

Relative ascidian cover = sum(ascidian cover) / sum(all other cover)
```{r}
colnames(benthic.data)
# make new data frame with columns without hard corals and sand (easier than taking them out of current df)
# also excluded dead coral and rubble
selected.cols <- c("TP", "Site", "Transect_ID", "Transect_no", "Anemone", "ASCD", "BInv", "CalcInvert", "SC", "SP", "Zoo", "Sediment", "Rhodo", "cyafil", "Cyanob", "TAS", "AL_fleshy", "AL_fol", "Amansia",  "ampart", "Asparago", "Bryopsis","calc_macro",  "Caulcupre", "CauRac", "CCA", "Chaeto", "Chlor", "CodFrag", "Colpo", "CYANO_DHC","Dasy","Dictyopter","Dictyota", "Halimeda","HALY", "Hormo", "Hydrocl", "Hyp", "Laur","MACR_Fil_A","MACR_LTH_O","Mal","MLRP", "Padina", "Ploc", "Sargassum", "Tri_fra", "Turf", "Ulva","Valspp" )
rel.cover.data <- benthic.data[selected.cols]

# calculate mean cover by transect -> transpose first
t.rel.cover <- rel.cover.data %>% pivot_longer(c(Anemone:Valspp), names_to = "Genera", values_to = "Cover")

rel.cover.mean.transects <- t.rel.cover %>%
  group_by(TP, Site, Transect_ID, Genera) %>%
  summarize(mean_cover = mean(Cover))

# transpose back
rel.cover.mean.transects <- rel.cover.mean.transects %>%
  pivot_wider(names_from = Genera, values_from = mean_cover)

# relative ascidian cover = sum(ascidian cover) / sum(all other cover)
rel.ascd.cover <-rel.cover.mean.transects %>%
  group_by(TP, Site, Transect_ID) %>%
  summarize(rel_ascd_cover = ASCD/sum(AL_fleshy:cyafil))

# if ASCD = 0 -> NaN
# if total sum = 0 -> Inf
```





Observation: where are the ascidians growing on? + close-up photographs? which "species"?


