---
title: "Ascidians 2021 and 2023"
author: "Shannon Eckhardt"
date: "2023-06-26"
output: html_document
---

This is the file for the report that looks at Ascidians in March 2021 and and April 2023, 1 and 3 years after the bleaching event in 2020.

```{r}
rm(list=ls())
library(tidyverse)
library(readxl)
library(dplyr)
library(stringr)
```

***2023***
**Import data**

RAW_EB_April2023 stems from the Master data - it contains over 15 transects for EB. All the transects that are above transect EB15 need to be removed because the 2021 data does not contain more than 15 transects for EB.

RAW_SB_April2023 stems from the Master data - it contains 25 transects for SB. All the transects that are above transect SB25 need to be removed because the 2021 data does not contain more than 25 transects for SB.
```{r}
raw_EB <- read_excel("Ascidian Data/RAW_EB_April2023.xlsx")
raw_SB <- read_excel("Ascidian Data/RAW_SB_April2023.xlsx")
char <- read_excel("Benthic_cover.xlsx")
```

**Data cleaning**
```{r}
#EB
# pivot the data frame to wide format
raw_EB_wide <- raw_EB %>% 
  group_by(Name, Label) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = Label, values_from = n, values_fill = 0)

# rename Name = Image_ID
# Pick out Transect_ID
# Pick out PQ = photoquadrat
# Extract TP, Transect_ID, and PQ from the Name column
raw_EB_wide <- raw_EB_wide %>%
  mutate(TP = "April_2023",
         Site = str_extract(Name, "EB"),
         Transect_ID = str_extract(Name, "EB\\d+"),
         Transect_no = str_extract(Name, "\\d+(?=\\.JPG$)")) %>%
  select(Name, TP, Site, Transect_ID, Transect_no, everything()) # reorder columns

# remove all the EB transects that are above 15
raw_EB_wide <- raw_EB_wide %>%
  filter(as.numeric(str_extract(Transect_ID, "\\d+")) <= 15)

#SB
raw_SB_wide <- raw_SB %>% 
  group_by(Name, Label) %>% 
  summarise(n = n()) %>% 
  pivot_wider(names_from = Label, values_from = n, values_fill = 0)

# rename Name = Image_ID
# Pick out Transect_ID
# Pick out PQ = photoquadrat
# Extract TP, Transect_ID, and PQ from the Name column
raw_SB_wide <- raw_SB_wide %>%
  mutate(TP = "April_2023",
         Site = str_extract(Name, "SB"),
         Transect_ID = str_extract(Name, "SB\\d+"),
         Transect_no = str_extract(Name, "\\d+(?=\\.JPG$)")) %>%
  select(Name, TP, Site, Transect_ID, Transect_no, everything()) # reorder columns

# remove all the SB transects that are above 25
raw_SB_wide <- raw_SB_wide %>%
  filter(as.numeric(str_extract(Transect_ID, "\\d+")) <= 25)
```

Extract EB01-15 and SB01-25 from Charlotte's data from 2023 and the 3 CB transects
!!! in the master data the formatting is EB1 and in Char's data the formatting is EB01
```{r}
char.2023 <- char[char$TP == "April_2023",]
# rename the transect IDs for them to be unique (from the master data transect IDs)
# CB can stay the same, just take the 0 out of the name
char.2023 <- char.2023 %>%
  mutate(Transect_ID = str_replace(Transect_ID, "CB0([1-3])", "CB\\1"))
# EB16-21 (6 additional transects)
# SB26-31 (6 additional transects)
char.2023 <- char.2023 %>%
  mutate(Transect_ID = case_when(
    Transect_ID == "EB0" ~ "EB16",
    Transect_ID == "EB02" ~ "EB17",
    Transect_ID == "EB05" ~ "EB18",
    Transect_ID == "EB06" ~ "EB19",
    Transect_ID == "EB07" ~ "EB20",
    Transect_ID == "EB11" ~ "EB21",
    Transect_ID == "SB01" ~ "SB26",
    Transect_ID == "SB02" ~ "SB27",
    Transect_ID == "SB03" ~ "SB28",
    Transect_ID == "SB04" ~ "SB29",
    Transect_ID == "SB05" ~ "SB30",
    Transect_ID == "SB06" ~ "SB31",
    TRUE ~ Transect_ID
  ))
```


Group everything into
- Macroalgae
- Turf
- Cyanobacteria
- Hard corals
- Other
- Ascidians
```{r}
# out EB and SB cleaned raw datasets together
EB_SB_2023 <- rbind(raw_EB_wide, raw_SB_wide)

#there's some NAs in data frame -> replace with 0 (those were not in labelset?)
EB_SB_2023 <- EB_SB_2023 %>%
  mutate(across(everything(), ~ifelse(is.na(.), 0, .)))

# transpose data
t.EB_SB_2023 <- EB_SB_2023 %>% pivot_longer(c(ABH:PLD), names_to = "Genera", values_to = "Cover")

benthic.cats <- t.EB_SB_2023 %>%
  mutate(Category = 
           ifelse(Genera %in% c("ABB", "ABD", "ABH", "ABP", "ACBB", "ACDD", "ACH", "ACPP", "ANB", "AND", "ANH", "ANP", "GOB", "GOD", "GOH", "GOP", "MEB", "MEDD", "MEH", "MEP", "MPB", "MPD", "MPH","MPP", "OBBB", "OBD", "OBH", "OBP", "PLB", "PLD", "PLH", "PLP", "POB","POD","POH", "POP"), 
                  "Hard coral",
                  ifelse(Genera %in% c("M", "K", "F", "OG"),
                         "Macroalgae",
                         ifelse(Genera %in% "G",
                                "Other benthic invertebrate",
                                       ifelse(Genera %in% c("S", "BARK", "R", "Rubble"),
                                              "Sand_sediment",
                                                     ifelse(Genera %in% c("RT", "BT"),
                                                            "Cyanobacteria",
                                                                   ifelse(Genera %in% "ASCD",
                                                                          "Ascidian",
                                                                          ifelse(Genera %in% "NA",
                                                                                 "All other", 
                                                                                 ifelse(Genera %in% "T",
                                                                                        "Turf",
                                                                                        "Rest")))))))))

# Group Char's data
t.char.2023 <- char.2023 %>% pivot_longer(c(Acanthastrea:Valspp), names_to = "Genera", values_to = "Cover")

char.benthic.cats <- t.char.2023 %>%
  mutate(Category = 
    ifelse(Genera %in% c("Acanthastrea","Alveopora","Astrea","Euphyllia", "Paragoni","Pectinia","Platy", "Plesiastrea","Porites","Psammo","Cyphastrea","Homophyllia.sp.","Lobophyllia","Turbinaria","PoritPale","PortiPal", "Pocillopora","Styloph","Stylophora","B_Pocillo","B_Styloph","PPoc", "Acropora_br", "Acropora_pl","PACR_br","PAcr_tab", "Monti_encr","Monti_fol","B_Monti","B_MontiEnc","PEncr"),
             "Hard coral",
    ifelse(Genera %in% c("cyafil","Bryopsis","Chaeto","Chlor","MACR_Fil_A", "Turf", "TAS"),
           "Turf",
    ifelse(Genera %in% c("CYANO_DHC", "Cyanob"),
           "Cyanobacteria",
    ifelse(Genera %in% c("AL_fol","Dasy","Ulva", "AL_fleshy","Amansia","Asparago","CodFrag","Hyp","Laur","Mal","Sargassum", "Colpo","Hormo","Hydrocl","MACR_LTH_O","MLRP","Padina","Ploc","Valspp", "ampart","calc_macro","Halimeda","HALY","Tri_fra", "Rhodo","CCA", "Dictyopter","Dictyota", "Caulcupre","CauRac"),
           "Macroalgae",
    ifelse(Genera %in% c("SC","Anemone", "SP","Zoo","BInv","CalcInvert"), 
             "Other benthic invertebrate",
    ifelse(Genera %in% "ASCD", 
             "Ascidian",
    ifelse(Genera %in% c("Sand","Sediment", "D_coral", "Rubble"),
           "Sand_sediment",
           NA ))))))))

unique(char.benthic.cats$Category) # where have the ascidians gone??

```

Sum counts per broad category per quadrat (Transect_no)
```{r}
# EB SB master data
EB_SB_cats_cover_sum <- benthic.cats %>%
  group_by(TP, Site,Transect_ID,Transect_no,Category) %>% # no grouping by TP as it's given (2023) %>%
  summarize(sum_count = sum(Cover))

# Char's data
char_cats_cover_sum <- char.benthic.cats %>%
  group_by(TP, Site,Transect_ID,Transect_no,Category) %>% # no grouping by TP as it's given (2023) %>%
  summarize(sum_count = sum(Cover))
```
***Bring EB SB master data and Char's data together***
```{r}
# Transect_no in EB_SB_cats_cover_sum is character -> convert into numeric
EB_SB_cats_cover_sum$Transect_no <- as.numeric(EB_SB_cats_cover_sum$Transect_no)

full_2023 <- rbind(char_cats_cover_sum, EB_SB_cats_cover_sum)

unique(full_2023$Transect_ID)
```

***Calculate TRANSECT mean cover***
```{r}
transect.percent.cover <- full_2023 %>%
  group_by(TP, Site, Transect_ID, Category) %>%
  summarize(percent_cover = mean(sum_count)) %>% # mean() already gives a percentage
  mutate(SE = sd(percent_cover) / sqrt(n()))

site.percent.cover <- full_2023 %>%
  group_by(TP, Site, Category) %>%
  summarize(percent_cover = mean(sum_count)) %>%
  mutate(SE = sd(percent_cover) / sqrt(n()))

ggplot(transect.percent.cover, aes(x = Category, y = percent_cover, fill = Site)) +
  geom_boxplot() +
  stat_boxplot(geom="errorbar") +
  labs(x = "Benthic category", y = "Cover (%)") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# just the ascidians
ascd.transect.percent.cover <- transect.percent.cover[transect.percent.cover$Category == "Ascidian",]
ascd.site.percent.cover <- site.percent.cover[site.percent.cover$Category == "Ascidian",]

ggplot(ascd.transect.percent.cover, aes(x = Site, y = percent_cover, fill = Site)) +
  geom_boxplot() +
  stat_boxplot(geom="errorbar") +
  labs(x = "Site", y = "Cover (%)") +
  theme_classic()
  
# dotplot with mean PER SITE and errorbars
ggplot(ascd.site.percent.cover, aes(x = Site, y = percent_cover, colour = Site)) +
  geom_point(position = position_dodge(.9), size = 2) +
  geom_errorbar(aes(ymin = percent_cover - SE, ymax = percent_cover + SE),
                width = 0.4, size = 1, position = position_dodge(.9)) +
  labs(x = "Site", y = "Cover (%)") +
  ylim(0, max(ascd.site.percent.cover$percent_cover + ascd.site.percent.cover$SE)) +  # adjust ylim to include the error bars
  theme_classic()
```
***Relative cover***
- take out all hard coral substrate (and not sand)
- for anything that isn't coral (or sand), what proportion is made up of ascidians
- how much of 100% is ascidian cover
- sum of ascidians divided by sum all other cover (including ascidians) minus hard coral and sand
- are they making up more cover than algae in some places
```{r}
no.hard.coral <- full_2023[full_2023$Category != "Hard coral",]

# get a df with just the ascidian count PER transect -> so 1 count for each transect
ascidian_df <- no.hard.coral %>%
  group_by(TP, Site, Transect_ID) %>%
  filter(Category == "Ascidian") %>%
  summarize(ascidian.count = sum(sum_count))

everything_df <- no.hard.coral %>%
  group_by(TP, Site, Transect_ID) %>%
  summarize(everything.count = sum(sum_count))

# bind together
rel.cover <- left_join(ascidian_df, everything_df)

# calculate relative cover of ascidians
rel.cover.transect <- rel.cover %>%
  group_by(TP, Site, Transect_ID) %>%
  summarize(rel_cover = (ascidian.count/everything.count)*100)
  
rel.cover.site <- rel.cover %>%
  group_by(TP,Site) %>%
  summarize(rel_cover = (ascidian.count/everything.count)*100) %>%
  mutate(SE = sd(rel_cover) / sqrt(n()))

# plot relative cover
ggplot(rel.cover.transect, aes(x = Site, y = rel_cover, fill = Site)) +
  geom_boxplot() +
  stat_boxplot(geom = "errorbar") +
  labs(x = "Site", y = "Relative Cover (%)") +
  theme_classic()

ggplot(rel.cover.site, aes(x = Site, y = rel_cover, colour = Site)) +
  geom_point(position = position_dodge(.9), size = 2) +
  geom_errorbar(aes(ymin = rel_cover - SE, ymax = rel_cover + SE),
                width = 0.4, size = 1, position = position_dodge(.9)) +
  labs(x = "Site", y = "Cover (%)") +
  ylim(0, max(rel.cover.site$rel_cover + rel.cover.site$SE)) +  # adjust ylim to include the error bars
  theme_classic()

```



