---
title: "R Notebook"
output: html_notebook
---


```{r}
rm(list=ls())
library(tidyverse)
library(dplyr)
library(readxl)
```

```{r}
benthic.data <- read_excel("Benthic_cover.xlsx") # Charlotte's data with CB sites that I will use
ascidians.full <- benthic.data[, c("Image ID", "Image name", "TP", "Site", "Transect_ID", "Transect_no", "ASCD")]
master <- read_excel("master_data_updated_april_2023.xlsx") # Masterdata with all the replicated transects
```

Extract April 2023 and 2021 data (2023 - 3 years after bleaching, 2021 - 1 year after bleaching)
```{r}
benthic.data.2023 <- benthic.data[benthic.data$TP == "April_2023",]

sum(benthic.data.2023$ASCD) # 69 ASCD recorded in 2023

# long format to group
t.benthic.data.2023 <- benthic.data.2023 %>% pivot_longer(c(Acanthastrea:Valspp), names_to = "Genera", values_to = "Cover")

```

Extract April 2023 from Master to re-analyze EB and SB transects
In March 2021 there are these transects:
- EB1 - EB15
- SB1 - SB25
These transects and the photos therein will be re-analyzed in April 2023 to pick out the ascidians.
Therefore, EB will have 15 replicates, and SB will have 25 replicates in each year.
For CB, I will take the CB 2021 & 2023 (3 transects each) from Charlotte's data
```{r}
unique(master$TP)
april2023 <- master[master$TP == "April_2023", ]

t.april2023 <- april2023 %>% pivot_longer(c(`Acanthastrea Bleached`:Black_turf), names_to = "Genera", values_to = "Cover")

# How many quadrats (PQ) contain values > 0 for "All other"
filtered_quadrats <- t.april2023 %>%
  group_by(Transect_ID, PQ, Genera) %>%
  filter(Genera == "All other", Cover > 0) # 249 quadrats have "all other" annotations

length(unique(filtered_quadrats$Transect_ID)) # 71 transects have all other" annotations


# remove CEM
filtered_quadrats_noCEM <- filtered_quadrats[!grepl("CEM_", filtered_quadrats$Transect_ID), ] # western slaughter still included, but am not re-analyzing these pictures
```
There are a few quadrats in EB and SB that contain "all other" annotations. These quadrats will be re-analyzed and "all other" will be reclassified to find any ascidians. In the master dataset, these images are titled with "March", e.g. *March_2023EB13_1.JPG*

Afterwards, the data for April 2023 will be exported into an excel file and data analysis will be done on the transects EB1-15 and SB1-25. The same thing will later be done for March 2021, however the quadrats containing the ascidians will be different.













***Absence/presence in transects***
Presence/absence table of ascidians
```{r}
# group by transect
sum.count.2023 <- t.benthic.data.2023 %>%
  group_by(Site, Transect_ID, Genera) %>%
  summarize(sum_count = sum(Cover))

# back to wide format
sum.count.2023 <- sum.count.2023 %>%
  pivot_wider(names_from = Genera, values_from = sum_count)

# presence/absence of ascidians per row = per Transect_ID
ascd.occurr.2023 <- sum.count.2023 %>%
  mutate(ascd_occurrence = ifelse(ASCD != 0, "present", "absent")) # only present on CB02 and CB03
```

***Absence/presence in quadrats***
```{r}
quadat.image.id <- benthic.data.2023 %>%
  mutate(ascd_occurrence = ifelse(ASCD != 0, "present", "absent"))

# take all the rows out where ascd_occurrence == "present"
quadrat.image.id.present <- quadat.image.id[quadat.image.id$ascd_occurrence == "present",] # 8 quadrats
```

***Counts over time for the transects that have ascidians***
```{r}
sum(benthic.data$ASCD)
t.benthic.data <- benthic.data %>% pivot_longer(c(Acanthastrea:Valspp), names_to = "Genera", values_to = "Cover")

sum.count <- t.benthic.data %>%
  group_by(TP, Site, Genera) %>%
  summarize(sum_count = sum(Cover))

# calculate mean cover per transect
mean.cover <- t.benthic.data %>%
  group_by(TP, Site, Transect_ID, Genera) %>%
  summarize(mean_cover = mean(Cover)) %>%
  mutate(SE = sd(mean_cover) / sqrt(n()),
         SD = sd(mean_cover))

# calculate percent cover per site
site.percent.cover <- mean.cover %>%
  group_by(TP, Site) %>%
  summarize(percent_cover = sum(mean_cover[Genera == "ASCD"]) / sum(mean_cover) * 100, .groups = "drop") %>%
  mutate(SE = sd(percent_cover)/sqrt(n()))

# sort TPs
site.percent.cover <- site.percent.cover %>%
  mutate(TP = factor(TP, levels = c("Dec_2020", "April_2021", "April_2022", "Sept_2022", "Dec_2022",  "April_2023"))) %>%
  arrange(TP)



# back to wide format
sum.count <- sum.count %>%
  pivot_wider(names_from = Genera, values_from = sum_count)

# presence/absence of ascidians per row = per Transect_ID
ascd.occurr <- sum.count %>%
  mutate(ascd_occurrence = ifelse(ASCD != 0, "present", "absent")) # only present on CB02 and CB03

allyears.transect.present <- ascd.occurr[ascd.occurr$ascd_occurrence == "present",]



# calculate percent cover on quadrat level = Transect_no to make boxplots
quadrat.percent.cover <- t.benthic.data %>%
  group_by(TP, Site, Transect_ID, Transect_no) %>%
  summarize(percent_cover = sum(Cover[Genera == "ASCD"]) / sum(Cover) * 100, .groups = "drop")

# transect percent cover
transect.percent.cover <- mean.cover %>%
  group_by(TP, Site, Transect_ID) %>%
  summarize(percent_cover = sum(mean_cover[Genera == "ASCD"]) / sum(mean_cover) * 100, .groups = "drop")

# sort TPs
quadrat.percent.cover <- quadrat.percent.cover %>%
  mutate(TP = factor(TP, levels = c("Dec_2020", "April_2021", "April_2022", "Sept_2022", "Dec_2022",  "April_2023"))) %>%
  arrange(TP)

transect.percent.cover <- transect.percent.cover %>%
  mutate(TP = factor(TP, levels = c("Dec_2020", "April_2021", "April_2022", "Sept_2022", "Dec_2022",  "April_2023"))) %>%
  arrange(TP)



# boxplot based on quadrat level percent cover of ascidians
ggplot(quadrat.percent.cover, aes(x = TP, y = percent_cover, fill = Site)) +
  geom_boxplot() +
  labs(x = "Time Points", y = "Cover (%)") +
  theme_classic()

# boxplot based on transect level percent cover of ascidians
ggplot(transect.percent.cover, aes(x = TP, y = percent_cover, fill = Site)) +
  geom_boxplot() +
  labs(x = "Time Points", y = "Cover (%)") +
  theme_classic()

ggplot(site.percent.cover, aes(x = TP, y = percent_cover, color = Site)) +
  geom_point() +
  labs(x = "Time Points", y = "Cover (%)") +
  theme_classic()
```
***Frequency of occurrence per site***
FOC = number of transects where ascidians were present / number of total transects
e.g. there's 3 transects in total in CB, if there was ascidians present in 2 of them, then 2/3 = FOC

