---
title: "3 TP Ascidian Analysis"
author: "Shannon Eckhardt"
date: "2023-09-22"
output: html_document
---

# IMPORT LIBRARIES
```{r}
rm(list=ls())
library(tidyverse)
library(readxl)
library(openxlsx)
library(RColorBrewer)
```


# IMPORT ALL DATA
```{r}
full_data <- read_excel("/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/01 - Data/FULL_DATA_21_22_23.xlsx")
# interactions <- read_excel("Ascidian Data/Interactions.xlsx") DO THE INTERACTION DATA LATER BECAUSE IT NEEDS TO BE REDONE
```

# TRANSECT MEAN COVER
```{r}
transect.percent.cover <- full_data %>%
  group_by(TP, Site, Transect_ID, Category) %>%
  summarize(percent_cover = mean(sum_count)) %>% # mean() already gives a percentage
  mutate(SE = sd(percent_cover) / sqrt(n()))

site.percent.cover <- full_data %>%
  group_by(TP, Site, Category) %>%
  summarize(percent_cover = mean(sum_count)) %>% # mean() already gives a percentage
  mutate(SE = sd(percent_cover) / sqrt(n()))

tp.percent.cover <- full_data %>%
  group_by(Site, Category) %>%
  summarize(percent_cover = mean(sum_count)) %>% # mean() already gives a percentage
  mutate(SE = sd(percent_cover) / sqrt(n()))
```

# EXTRACT ASCIDIAN DATA
```{r}
# just the ascidians
ascd.transect.percent.cover <- transect.percent.cover[transect.percent.cover$Category == "Ascidians",]

# one df per site
SB_ascd_cover <- ascd.transect.percent.cover[ascd.transect.percent.cover$Site == "SB",]
EB_ascd_cover <- ascd.transect.percent.cover[ascd.transect.percent.cover$Site == "EB",]
CB_ascd_cover <- ascd.transect.percent.cover[ascd.transect.percent.cover$Site == "CB",]


#relevel sites, SB, EB, CB
ascd.transect.percent.cover$Site <- factor(ascd.transect.percent.cover$Site, levels = c("SB", "EB", "CB"))
```
# EXTRACT SAND DATA
```{r}
# sand percent cover
sand.percent.cover <- transect.percent.cover[transect.percent.cover$Category == "Sand and sediment",]
# rename percent_cover to sand_percent_cover
sand.percent.cover <- sand.percent.cover %>% rename(sand_percent_cover = percent_cover)
# add sand percent cover column to ascidian data to make new df which includes cover for ascidians and sand

ascd.sand.cover <- left_join(ascd.transect.percent.cover, 
                       sand.percent.cover %>% dplyr::select(TP, Site, Transect_ID, sand_percent_cover),
                       by = c("TP", "Site", "Transect_ID"))

# add 
```


# FREQUENCY OF OCCURRENCE (FOC)
```{r}
# filter where ASCD cover > 0
foc <- ascd.transect.percent.cover %>%
  filter(percent_cover > 0)

# group by Site and count the number of unique Transect_IDs with sum_count of Ascidian greater than 1
foc <- foc %>%
  group_by(TP,Site) %>%
  summarise(Num_Transects_with_Ascidian = n_distinct(Transect_ID))

```

# ASCIDIAN COVER PER SITE
```{r}
# pick site colors that match map in manuscript methods
site_colors <- c("CB" = "#FE7F9C", "EB" = "#EC9706", "SB" = "#52B2BF")

# SB: one outlier in 2023:
  # SB12 -> percent cover of 4.9%
  # removing from plot and then mention in figure legend "outlier in 2023 was removed with a value of 4.9%
#SB_ascd_cover <- subset(SB_ascd_cover, Transect_ID != "SB12")

SBcover <- ggplot(SB_ascd_cover, aes(x = TP, y = percent_cover, fill = Site)) +
  geom_boxplot(width = 0.3) +
  #stat_boxplot(geom = "errorbar", width = 0.3) +
  labs(x = "Time point", y = "Cover (%)") +
  scale_fill_manual(values = site_colors) +
  theme_classic() +
  #ylim(0,1) +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    legend.position = "bottom",  # Set legend position to bottom
    legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(1.2, "cm"))  # Adjust the size of the legend key

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/SB_ascidian_cover.jpg", plot = SBcover, width = 6, height = 5, dpi = 300)


EBcover <- ggplot(EB_ascd_cover, aes(x = TP, y = percent_cover, fill = Site)) +
  geom_boxplot(width = 0.3) +
  #stat_boxplot(geom = "errorbar", width = 0.3) +
  labs(x = "Time point", y = "Cover (%)") +
  scale_fill_manual(values = site_colors) +
  theme_classic() +
  #ylim(0,1) +
  theme(
    text = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size =18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = "bottom",
    legend.justification = "center",
    legend.key.size = unit(1.2, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/EB_ascidian_cover.jpg", plot = EBcover, width = 6, height = 5, dpi = 300)


CBcover <- ggplot(CB_ascd_cover, aes(x = TP, y = percent_cover, fill = Site)) +
  geom_boxplot(width = 0.3) +
  #stat_boxplot(geom = "errorbar", width = 0.3) +
  labs(x = "Time point", y = "Cover (%)") +
  scale_fill_manual(values = site_colors) +
  theme_classic() +
  #ylim(0,5) +
  theme(
    text = element_text(size = 16),
    axis.text.x = element_text(size = 16),
    axis.text.y = element_text(size = 16),
    axis.title.x = element_text(size = 18),
    axis.title.y = element_text(size =18),
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18),
    legend.position = "bottom",
    legend.justification = "center",
    legend.key.size = unit(1.2, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/CB_ascidian_cover.jpg", plot = CBcover, width = 6, height = 5, dpi = 300)


```

# MEAN COVER ALL CATEGORIES
```{r}
CB.site.percent.cover <- site.percent.cover[site.percent.cover$Site == "CB",]
SB.site.percent.cover <- site.percent.cover[site.percent.cover$Site == "SB",]
EB.site.percent.cover <- site.percent.cover[site.percent.cover$Site == "EB",]

write.xlsx(site.percent.cover, file = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/04 - Tables/Cover_averaged_over_sites.xlsx", rowNames = FALSE)


CB.cat.cov <- ggplot(CB.site.percent.cover, aes(x = TP, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/CB_category_cover.jpg", plot = CB.cat.cov, width = 8, height = 5, dpi = 300)

SB.cat.cov <- ggplot(SB.site.percent.cover, aes(x = TP, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/SB_category_cover.jpg", plot = SB.cat.cov, width = 8, height = 5, dpi = 300)

EB.cat.cov <- ggplot(EB.site.percent.cover, aes(x = TP, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))
  #guides(fill = guide_legend(nrow = 5))  # Split the legend into 2 rows

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/EB_category_cover.jpg", plot = EB.cat.cov, width = 8, height = 5, dpi = 300)

```

# MEAN COVER PER SITE OVER THREE TIME POINTS
```{r}
CB.tp.percent.cover <- tp.percent.cover[tp.percent.cover$Site == "CB",]
SB.tp.percent.cover <- tp.percent.cover[tp.percent.cover$Site == "SB",]
EB.tp.percent.cover <- tp.percent.cover[tp.percent.cover$Site == "EB",]

tp.percent.cover$Site <- factor(tp.percent.cover$Site, levels = c("SB", "EB", "CB"))

write.xlsx(tp.percent.cover, file = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/04 - Tables/Cover_averaged_over_3TP.xlsx", rowNames = FALSE)


# cover of all sites in one plot
all.tp.cov <- ggplot(tp.percent.cover, aes(x = Site, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16), 
    axis.text.x = element_text(size = 16),  
    axis.text.y = element_text(size = 16),  
    axis.title.x = element_text(size = 18),  
    axis.title.y = element_text(size =18), 
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/All_site_cover.jpg", plot = all.tp.cov, width = 8, height = 5, dpi = 300)


CB.tp.cov <- ggplot(CB.tp.percent.cover, aes(x = Site, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))
  #guides(fill = guide_legend(nrow = 4))  # Split the legend into 2 rows

SB.tp.cov <- ggplot(SB.tp.percent.cover, aes(x = Site, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16), 
    axis.text.x = element_text(size = 16),  
    axis.text.y = element_text(size = 16),  
    axis.title.x = element_text(size = 18),  
    axis.title.y = element_text(size =18), 
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

EB.tp.cov <- ggplot(EB.tp.percent.cover, aes(x = Site, y = percent_cover, fill = Category)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_brewer(palette = "BrBG") +  
  theme_classic() +
  theme(
    text = element_text(size = 16), 
    axis.text.x = element_text(size = 16),  
    axis.text.y = element_text(size = 16),  
    axis.title.x = element_text(size = 18),  
    axis.title.y = element_text(size =18), 
    legend.text = element_text(size = 16),
    legend.title = element_text(size = 18), 
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/CB_site_cover.jpg", plot = CB.tp.cov, width = 6, height = 5, dpi = 300)
ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/SB_site_cover.jpg", plot = SB.tp.cov, width = 6, height = 5, dpi = 300)
ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/EB_site_cover.jpg", plot = EB.tp.cov, width = 6, height = 5, dpi = 300)

```


# STATISTICS FOR COVER
ascd.transect.percent.cover.stats includes all transects, including those that do not contain ascidians
SB: 23
EB: 15
CB: 5
```{r}
ascd.transect.percent.cover.stats <- ascd.transect.percent.cover %>% 
  mutate(log_percent_cover = log(percent_cover+1))

# look at distribution of percent_cover
ggplot(ascd.transect.percent.cover.stats, aes(x = percent_cover)) +
  geom_histogram()

ggplot(ascd.transect.percent.cover.stats, aes(x = log_percent_cover)) +
  geom_histogram()


m1 <- lm(log_percent_cover ~ Site * TP, data = ascd.transect.percent.cover.stats)

plot(m1)
anova(m1)
summary(m1)
  # significant effect of interaction term p < 0.01
  # from summary plot, get R squared, nothing more

# post hoc testing
library(emmeans)

posthoc <- lsmeans(m1, pairwise ~ Site * TP, adjust = "Tukey") # adjusted using tukey-ascombe?


#lsmeans_df <- data.frame(posthoc$contrasts)
#write.xlsx(lsmeans_df, file = "Ascidian Data/lsmeans_results.xlsx", rowNames = FALSE)


# No log-transformation
m2 <- lm(percent_cover ~ Site * TP, data = ascd.transect.percent.cover)
plot(m2)
anova(m2)
summary(m2)
lsmeans(m2, pairwise ~ Site * TP, adjust = "Tukey")


```

## BETA BINOMIAL DISTRIBUTION
```{r}
library(glmmTMB)

# convert percent cover from 0-100 to 0-1
ascd.transect.percent.cover$percent_cover <- ascd.transect.percent.cover$percent_cover / 100

m.bb <- glmmTMB(percent_cover ~ TP * Site + (1|Transect_ID),
                 family = betabinomial(link = "logit"),
                 data = ascd.transect.percent.cover)

diagnose()


```

## HURDLE MODEL
```{r}
m.hurd <- glmmTMB(percent_cover ~ TP * Site + (1|Transect_ID), ziformula = ~1, data = ascd.transect.percent.cover, family=beta_family())
m.hurd.sand <- glmmTMB(percent_cover ~ TP * Site+ sand_percent_cover + (1|Transect_ID), ziformula = ~1, data = ascd.sand.cover, family=beta_family())



library(DHARMa)

simulationOutput <- simulateResiduals(fittedModel = m.hurd, plot = F)
simulationOutput.sand <- simulateResiduals(fittedModel = m.hurd.sand, plot = F)
plot(simulationOutput)
plot(simulationOutput.sand)



library(car)
Anova(m.hurd)
Anova(m.hurd.sand)

library(emmeans)
emmeans(m.hurd, ~ TP * Site, component = "response")
?emmeans

```






# INTERACTIONS
```{r}
interactions <- read_excel("/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/01 - Data/Interactions.xlsx")
```

## SUBSTRATE SINGLE TP
```{r}
# Define the custom colors for each substrate type
substrate_types <- c("Cyanobacteria",
                     "Hard coral", 
                     "Hard substrate",
                     "Macroalgae",
                     "Sand and sediment",
                     "Turf")

substrate_colors <- brewer.pal(6, "BrBG")

#library(scales)
#substrate_colors_light <- alpha(substrate_colors, alpha = 0.75) # make it lighter/more transparent
substrate_color_map <- setNames(substrate_colors, substrate_types)


# create percentage
substrate.sum <- interactions %>%
  group_by(Site, Substrate) %>%
  summarize(sum_count = sum(n()))

total_counts <- substrate.sum %>%
  group_by(Site) %>%
  summarize(total_count = sum(sum_count))
 # total count of substrates/images where ascidians occur per TP

substrate.sum <- left_join(substrate.sum, total_counts, by = "Site")

# calculate the percentage of times ascidians occur on each substrate at each time point
substrate <- substrate.sum %>%
  mutate(percentage = (sum_count / total_count) * 100)

substrate$Site <- factor(substrate$Site, levels = c("SB", "EB", "CB"))

SUB <- ggplot(substrate, aes(x = Site, y = percentage, fill = Substrate)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_manual(values = substrate_color_map) +
  theme_classic() +
  ggtitle("Substrate") +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/Substrate_1TP.jpg", plot = SUB, width = 6, height = 5, dpi = 300)
write.xlsx(substrate, file = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/04 - Tables/Substrate_results.xlsx", rowNames = FALSE)


# overall substrate numbers

```

## INTERACTIONS
```{r}
# create percentage
interactions.sum <- interactions %>%
  group_by(Site, Interaction) %>%
  summarize(sum_count = sum(n()))

total_counts <- interactions.sum %>%
  group_by(Site) %>%
  summarize(total_count = sum(sum_count))
 # total count of substrates/images where ascidians occur per TP

interactions.sum <- left_join(interactions.sum, total_counts, by = "Site")

# calculate the percentage of times ascidians occur on each substrate at each time point
int <- interactions.sum %>%
  mutate(percentage = (sum_count / total_count) * 100)

int$Site <- factor(int$Site, levels = c("SB", "EB", "CB"))

INT <- ggplot(int, aes(x = Site, y = percentage, fill = Interaction)) +
  geom_bar(stat = "identity", position = "stack") +
  labs(y = "Frequency",
       x = "",
       fill = "") +
  scale_fill_manual(values = substrate_color_map) +
  theme_classic() +
  ggtitle("Interactions") +
  theme(
    text = element_text(size = 16),  # Adjust the overall font size
    axis.text.x = element_text(size = 16),  # Adjust x-axis tick label font size
    axis.text.y = element_text(size = 16),  # Adjust y-axis tick label font size
    axis.title.x = element_text(size = 18),  # Adjust x-axis label font size
    axis.title.y = element_text(size =18),  # Adjust y-axis label font size
    legend.text = element_text(size = 16),  # Adjust legend text font size
    legend.title = element_text(size = 18),  # Adjust legend title font size
    #legend.position = "bottom",  # Set legend position to bottom
    #legend.justification = "center",  # Set legend justification to center
    legend.key.size = unit(0.7, "cm"))

ggsave(filename = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/03 - Plots/Interactions_1TP.jpg", plot = INT, width = 6, height = 5, dpi = 300)
write.xlsx(int, file = "/Users/Shannon/GitHub/NFI_benthic_community/3 TP Analysis/04 - Tables/Interaction_results.xlsx", rowNames = FALSE)

```

## OVERALL SUBSTRATE AND INTERACTIONS NUMBERS
Overall (all sites)
```{r}
# SUBSTRATE
# create percentage
overall.substrate.sum <- interactions %>%
  group_by(Substrate) %>%
  summarize(sum_count = sum(n()))

total_counts <- overall.substrate.sum %>%
  summarize(total_count = sum(sum_count))
 # total count of substrates/images where ascidians occur per TP

# calculate the percentage of times ascidians occur on each substrate at each time point
overall.substrate <- overall.substrate.sum %>%
  mutate(percentage = (sum_count / 56) * 100)


# INTERACTIONS
overall.interactions.sum <- interactions %>%
  group_by(Interaction) %>%
  summarize(sum_count = sum(n()))

total_counts <- overall.interactions.sum %>%
  summarize(total_count = sum(sum_count))

# calculate the percentage of times ascidians occur on each substrate at each time point
overall.int <- overall.interactions.sum %>%
  mutate(percentage = (sum_count / 56) * 100)

```



